The fraction of polar aligned circumbinary disks

### 2025-04-29 src/tex/response.tex

\begin{letter}{}
    \opening{Dear Editor Carraro, Data Editors, and Anonymous Reviewer,}

    Please find our revised manuscript alongside the below response to the reviewer's report. The authors would like to thank the editor, reviewer, and data editors for their thoughtful review, which has lead to a greatly improved manuscript. Changes have been highlighted in \textcolor{blue}{\bf blue}.

    In addition to responding to the reviewer's report, the authors have made the following additional changes:

    \begin{enumerate}
        \item We have added text referencing the recent strong evidence for the first polar circumbinary planet\footnote{
            Baycroft, T. A., Sairam, L., Triaud, A. H. M. J., \& Correia, 760
        A. C. M. 2025, Science Advances, 11, eadu0627,761 doi: \href{http://doi.org/10.1126/sciadv.adu0627}{10.1126/sciadv.adu0627}}. Please see lines 41-46 and 739-742.
        \item We have added clarifying language to distinguish prograde/retrograde orbit of the disk from the retrograde/prograde nodal precession the disk undergoes. This includes changing the legend labels in many of the figures in an effort to reduce the potential for confusion. We more frequently refer to circulation rather than just stating that the disk is prograde/retrograde, and where appropriate we simply give the end state of the disk evolution (prograde coplanar, retrograde coplanar, polar).
    \end{enumerate}

@@

    \begin{response}
        {Fig. 8: The bottom part of this figure could be further commented in Sect. 6.1.}
        We have added text in Section 6.1 discussing this panel.
    \end{response}

    \begin{response}{
        L532: Bohn+2022 could be a relevant reference to add here.
    }
     Thank you for the great suggestion. We have added the reference.   
    \end{response}

    \begin{response}
        {L580: I would conclude the discussion (or add to the conclusion) that, besides imposing constraints on the disc radial extend and j, the two studied mechanisms (disc breaking and self-gravity) are expected to lower the fraction of polar CBDs. If I understood correctly, large values of j and radially extended discs are unlikely to become polar (as a whole at least). Therefore, the statistical relevance/weight of this subgroup of discs should be less important at the overall polar CBD population level. This could be briefly commented at the end.}

        This is a great suggestion, however the overall effect is more complicated. The main effect of both mechanisms is to shorten the alignment timescale so that fewer disks are left in an arbitrarily misaligned state at the end of their lives. We have added text explaining how this process works for gravitationally unstable disks and commenting on the shortened timescale for both effects. Additionally, we have made a change to the abstract and conclusion noting that effects due to non-zero radial extent impact the evolutionary timescale.
    \end{response}

### 2025-05-01 src/tex/response.tex

\begin{letter}{}
    \opening{Dear Editor Carraro, Data Editors, and Anonymous Reviewer,}

    Please find our revised manuscript alongside the below response to the reviewer's report. The authors would like to thank the editor, reviewer, and data editors for their thoughtful review, which has lead to a greatly improved manuscript. Changes have been highlighted in \textcolor{blue}{\bf blue}.

    In addition to responding to the reviewer's report, the authors have made the following additional changes:

    \begin{enumerate}
        \item We have added text referencing the recent strong evidence for the first polar circumbinary planet\footnote{
            Baycroft, T. A., Sairam, L., Triaud, A. H. M. J., \& Correia, 760
        A. C. M. 2025, Science Advances, 11, eadu0627,761 doi: \href{http://doi.org/10.1126/sciadv.adu0627}{10.1126/sciadv.adu0627}}. Please see lines 41-46 and 718-721.
        \item We have added clarifying language to distinguish the prograde/retrograde orbit of the disk from the retrograde/prograde nodal precession the disk undergoes. This includes changing the legend labels in many of the figures in an effort to reduce the potential for confusion. We more frequently refer to circulation rather than just stating that the disk is prograde/retrograde, and where appropriate we simply give the end state of the disk evolution (prograde coplanar, retrograde coplanar, polar).
    \end{enumerate}

### 2025-05-01 src/tex/ms.tex
Final edits

\newcommand{\TJ}[1]{\textcolor{orange}{#1}}
\newcommand{\RGM}[1]{\textcolor{cyan}{#1}}
\newcommand{\RR}[1]{\textcolor{blue}{\bf#1}} % reviewer's report
\newcommand{\RP}[1]{\textcolor{red}{\bf#1}} % Retrograde/Prograde confusion

@@

    Circumbinary gas disks that are misaligned to the binary orbital plane evolve toward either a coplanar or a polar-aligned configuration with respect to the binary host. The preferred alignment depends on the dynamics of the disk: whether it undergoes libration or circulating nodal precession, with librating disks evolving to polar inclinations and circulating disks evolving to coplanar.
    We quantify the fraction of binary star systems whose disks are expected to have polar orbits $f_\text{polar}$, extending previous work to include disks with non-zero mass.
    Our results suggest that, for low mass disks, the polar fraction is highly sensitive to the distribution of binary eccentricity with a higher fraction expected for higher binary eccentricities, $f_{\rm polar}\sim e_{\rm b}$. However, for massive discs, the fraction is independent of the binary eccentricity and $f_{\rm polar}\approx 0.37$. The value of $f_\text{polar}$ is always reduced in a population with a greater preference for low initial mutual inclination.
    We also explore the consequences of the finite lifetime and non-zero radial extent of a real disk, \RR{both of which affect a disk's ability to complete its evolution to a stationary configuration.} Our findings can be used to make predictions given populations with well-understood distributions of binary eccentricity, initial mutual inclination, and disk angular momentum.
\end{abstract}

@@

Given enough time, a gas disk that is misaligned from one of the two stationary inclinations moves to alignment through torques excited by warps that form as the disk undergoes differential precession \citep{bate2000,lubow2000}. Disks which are misaligned from coplanar only slightly \RR{(or slightly misaligned from retrograde coplanar $i\sim\pi$) undergo  retrograde (or prograde)} nodal precession \citep{papaloizou1995}. \citet{bate2000} and \citet{lubow2000} show that this kind of motion leads to coplanar alignment on a timescale that scales as $h^2/\alpha$, where $h$ and $\alpha$ are the disk aspect ratio and viscosity parameter, respectively.

When the mutual inclination is $\sim\pi/2$ (near polar), the disk can librate about the eccentricity vector of the binary \citep[e.g.,][]{verrier2009,farago2010,doolin2011}. In sufficiently massive disks, a third kind of motion arises at high inclinations. \RP{Retrograde nodal precession takes over from prograde nodal precession near the anti-aligned point}, and {\it crescent} orbits take over at inclinations between libration and high-inclination precession (\citealt{chen2019}, see also \citealt{zanazzi2018}), named because of the shape they form on a phase diagram. \citet{abod2022} showed that both librating and crescent orbiting disks decay to the stationary polar inclination \citep[see][]{martin2019} on an evolutionary timescale similar to that of coplanar alignment.

Thus, when the disk lifetime is longer than the alignment timescale, it is appropriate to map the secular dynamics (which we call the dynamical state, e.g. nodal precession, libration or crescent orbits) to an end state of the disk (coplanar or polar, respectively). More challenging is mapping the parameters of the disk (e.g., binary eccentricity, initial inclination, disk mass, etc.) to a dynamical state. In general, this can be done using expensive 3D disk simulations (e.g., those shown in \RR{\citealt{aly2015}, \citealt{cuello2019}} \citealt{martin2019}, \citealt{smallwood2019}, \citealt{abod2022}, \RR{\citealt{young2023} \& \citealt{ceppi2023}}). It is not feasible to perform such simulations over a large parameter space, so we instead approximate a disk as a geometrically narrow ring with angular momentum equivalent to that of the extended disk it represents. The literature provides a robust theoretical framework for the dynamics of such a ring, with analytical solutions \citep{farago2010,zanazzi2018,martin2019} as well as numerical $n$-body simulations \citep{abod2022,lepp2022} used to translate initial conditions into the ring's dynamical state. For simplicity, we assume that the narrow ring model approximates an extended disk sufficiently well. This is justified if warps and breaks are not important to the disk's evolution, though we discuss these effects in the context of our model.

\RP{
    For clarity and consistency with existing literature, it is important to note that disks which undergo retrograde nodal precession (those that are only slightly misaligned) also orbit the binary in the prograde direction, and herein we refer to them as {\em prograde circulating orbits}, or simply as {\em prograde}. Likewise, those disks which undergo prograde nodal precession ($i\sim\pi$) orbit their binary host in the retrograde direction, and we refer to them as {\em retrograde circulating orbits} or simply {\em retrograde}. As noted previously, the nodal precession direction for retrograde circulating orbits switches for sufficiently massive disks. We call these disks prograde after the transition to be consistent with their direction on the $(i,\,\Omega)$ phase diagram, but as all nodally precessing circumbinary disks will evolve towards a coplanar orientation, this distinction is not important.
}

@@

For an isotropic \RR{distribution of the angular momentum vector direction (i.e. one where the specific angular momentum $\bm{\ell}$ has no preference for any direction)}, the polar fraction in this limit is
\begin{align}
    f_{{\rm polar}, j \gg 1} &= \int_0^\pi\,\left (1-\frac{2}{\pi}\arcsin{\left (\sqrt{\frac{2}{5}} \csc{i} \right )}\right ) \sin{i}\,di\, , \\
    & = \variable{output/high_j_integral.txt} \, . \nonumber
\end{align}
This is the value that the higher $j$ curves in Figure \ref{fig:ebin} tend to for low $e_{\rm b}$, computed numerically.

@@

We determine the dynamical state of the system by tracking the orbital quantities $x=i\,\cos{\Omega}$ and $y=i\,\sin{\Omega}$, where
\begin{equation}
    i = \arccos{\ell_z}
    \label{eq:inclination}
\end{equation}
and
\begin{equation}
    \Omega = \arctantwo (\ell_x, -\ell_y)\, .
    \label{eq:omega}
\end{equation}
If a disk undergoes \RP{retrograde} nodal precession, then \RR{the path $(x,\,y)$} forms clockwise loops about the origin as the system evolves. Under \RP{prograde} precession, these loops are counter-clockwise. However, any state that would lead to a polar orbit (i.e. libration and crescent patterns) will form closed loops that do not encompass the origin, and these tracks will never cross the line $i\,\sin{\Omega}=0$. The state can be determined by tracking the quadrant \RR{(defined using the standard counter-clockwise convention)} of the point $(i\,\cos{\Omega},~i\,\sin{\Omega})$. \RP{Circulating orbits will traverse all four quadrants.}
A polar orbit that starts in quadrant I or II, however, is restricted to those two quadrants and will alternate between them. The same goes for polar orbits in quadrants III and IV. Therefore, the dynamical state can often be determined by the list of previous quadrant crossings. For example, I$\rightarrow$IV can only be caused by prograde \RP{circulation}, III$\rightarrow$IV is ambiguous, and III$\rightarrow$IV$\rightarrow$III indicates a polar orbit.

@@

For any given system, the dynamical state is determined similarly to the method described in Section \ref{sec:rk}. Note, however, that the $(i\,\cos{\Omega},~i\,\sin{\Omega})$ tracks now contain small structure on the orbital timescale of the outer mass. These structures can mean that a simulation may not make monotonic progress along its track, occasionally backtracking. Such a backtrack can give a false reading of the dynamical state if it occurs across the $x$ or $y$ axis. To mitigate this problem, we integrate each system until it returns to its original position in $(i\,\cos{\Omega},~i\,\sin{\Omega})$ space. If, at this point, the path does not enclose the origin, then it is polar. Prograde and retrograde circulating orbits can be distinguished by the sign of $\langle d\Omega/dt \rangle$. Figure \ref{fig:sim_example} shows paths for a set of simulations initialized with $\Omega = \pi/2$.

@@

\begin{figure}
    \begin{centering}
        \includegraphics[width=0.5\textwidth]{figures/compare_massive.pdf}
        \caption{
            MC results (points) for $e_{\rm b}=0.4$, $j=0.66$, and $f_{\rm b}=0.5$ compared to RKF results for the same setup (contour lines). This value of $j$ is chosen here as it is the critical value \citep[see][]{martin2019,abod2022}, beyond which \RP{prograde nodal precession is not allowed}. The regions interior to the black contour lines are initial conditions which allow polar orbits according to the RKF method.
            Note that some points that {\sc rebound} classifies as polar fall outside the RKF polar region.
            The MC result gives $f_{\text{polar}} = \variable{output/mc_out.txt}$ while the RKF result gives $f_{\text{polar}} = \variable{output/rk_out.txt}$ with
            numerical error on the order of $\variable{output/rk_err.txt}$. \RR{The MC simulation results used to generate this figure can be downloaded at \dataset[doi: 10.5281/zenodo.14736363]{\doi{10.5281/zenodo.14736363}}}
        }
        \script{compare_massive.py}
        \label{fig:rkf}
    \end{centering}
\end{figure}

@@

Figure \ref{fig:fpol_j} shows $f_\text{polar}$ as a function of $j$ for various choices of $p_{e_\text{b}}(e_\text{b})$. As expected, the curves all converge in the Kozai-Lidov regime. The uniform and thermal distributions both produce polar fractions $>0.5$ for small $j$, with the thermal distribution slightly higher. Except for when $j$ is close to 0, the circularized $e_\text{b} < 0.1$ curve follows the $e_\text{b} = 0.01$ curve very closely, including a maximum at $j=0.5$. This feature is present to some degree in all the low-$j$ curves, and is the result of the fact that, for $e_\text{b}=0$, the critical value for $j$ is 0.5 \citep{abod2022}. \RP{It is at this value of $j$ which crescent orbits begin to appear, and the size and shape of the libration region changes very rapidly. The transition in the nodal precession direction for retrograde circulating orbits also occurs here. For general $e_\text{b}$, the transition criterion is $j>j_\text{cr}$, where $j_\text{cr}$ is the critical value of $j$ \citep[eq. A3]{martin2019}.}

@@

Note that $\Gamma_1$ depends on $e_\text{b}$, and so it is only a constant in the case of a test particle. For $j>0$, we must compute the precession timescale ourselves. The time coordinate used in Section \ref{sec:rk} is $\tau \equiv \alpha' t$. $\alpha'$ is defined by \citet[see eq. 3.9]{farago2010}, and we write it as
\begin{equation}
    \alpha' \equiv \frac{3}{4} \Omega_\text{b} \left(\frac{a_\text{b}}{r}\right)^{7/2} f(1-f) \,\Gamma_2(j,f,a_\text{b}/r, e_\text{b}) \, ,
\end{equation}
where we have written the $\sqrt{1+m_\text{r}/M_\text{b}}$ factor as $\Gamma_2$. \RR{$m_\text{r}$ is the mass of the ring. To compute $\Gamma_2$ from our variables requires finding the roots of a cubic equation (see equation (\ref{eq:gamma2-eq})). However, for low $j$ (e.g. $j<10$), $m_\text{r} \ll M_\text{b}$ and $\Gamma_2 \sim 1$.}

>>>>

\newcommand{\TJ}[1]{\textcolor{orange}{#1}}
\newcommand{\RGM}[1]{\textcolor{cyan}{#1}}
\newcommand{\RR}[1]{\textcolor{blue}{\bf#1}} % reviewer's report
\newcommand{\RP}[1]{\RR{#1}} % Retrograde/Prograde confusion

@@

    Circumbinary gas disks that are misaligned to the binary orbital plane evolve toward either a coplanar or a polar-aligned configuration with respect to the binary host. The preferred alignment depends on the dynamics of the disk: whether it undergoes librating or circulating nodal precession, with librating disks evolving to polar inclinations and circulating disks evolving to coplanar.
    We quantify the fraction of binary star systems whose disks are expected to have polar orbits $f_\text{polar}$, extending previous work to include disks with non-zero mass.
    Our results suggest that, for low mass disks, the polar fraction is highly sensitive to the distribution of binary eccentricity with a higher fraction expected for higher binary eccentricities, $f_{\rm polar}\sim e_{\rm b}$. However, for massive discs, the fraction is independent of the binary eccentricity and $f_{\rm polar}\approx 0.37$. The value of $f_\text{polar}$ is always reduced in a population with a greater preference for low initial mutual inclination.
    We also explore the consequences of the finite lifetime and non-zero radial extent of a real disk, \RR{both of which affect a disk's ability to complete its evolution to a stationary configuration.} Our findings can be used to make predictions given populations with well-understood distributions of binary eccentricity, initial mutual inclination, and disk angular momentum.
\end{abstract}

@@

% Given enough time, a gas disk that is misaligned from one of the two stationary inclinations moves to alignment through torques excited by warps that form as the disk undergoes differential precession \citep{bate2000,lubow2000}. Disks which are misaligned from coplanar only slightly \RR{(or slightly misaligned from retrograde coplanar $i\sim\pi$) undergo  retrograde (or prograde)} nodal precession \citep{papaloizou1995}. \citet{bate2000} and \citet{lubow2000} show that this kind of motion leads to coplanar alignment \RGM{(or retrograde coplanar alignment)} on a timescale that scales as $h^2/\alpha$, where $h$ and $\alpha$ are the disk aspect ratio and viscosity parameter, respectively. 
\RR{Disks which are slightly misaligned from coplanar (or retrograde coplanar)  undergo circulating nodal precession about the binary angular momentum vector \citep{papaloizou1995}. \citet{bate2000} and \citet{lubow2000} show that this kind of motion leads to coplanar alignment {(or retrograde coplanar alignment)} on a timescale that scales as $h^2/\alpha$, where $h$ and $\alpha$ are the disk aspect ratio and viscosity parameter, respectively. Note that we use the terms prograde and retrograde to describe the direction of the orbit of the third body relative to the binary orbit, rather than the direction of the nodal precession. }

When the mutual inclination is $\sim\pi/2$ (near polar), the disk can librate about the eccentricity vector of the binary \citep[e.g.,][]{verrier2009,farago2010,doolin2011}. In \RR{addition, for} sufficiently massive disks, a third kind of motion arises at high inclinations. 
% \RP{Retrograde nodal precession takes over from prograde nodal precession near the anti-aligned point}, and {\it crescent} orbits take over at inclinations between libration and high-inclination precession (\citealt{chen2019}, see also \citealt{zanazzi2018}), named because of the shape they form on a phase diagram.
\RR{{\it Crescent} orbits take over at inclinations between libration and retrograde circulation, named because of the shape they form on a phase diagram  (\citealt{chen2019}, see also \citealt{zanazzi2018}). }
\citet{abod2022} showed that both librating and crescent orbiting disks decay to the stationary polar inclination \citep[see][]{martin2019} on an evolutionary timescale similar to that of coplanar alignment.

Thus, when the disk lifetime is longer than the alignment timescale, it is appropriate to map the secular dynamics (which we call the dynamical state, e.g. nodal precession, libration or crescent orbits) to an end state of the disk (coplanar or polar, respectively). More challenging is mapping the parameters of the disk-binary system (e.g., binary eccentricity, initial inclination, disk mass, etc.) to a dynamical state. In general, this can be done using expensive 3D disk simulations (e.g., those shown in \RR{\citealt{aly2015}, \citealt{cuello2019}} \citealt{martin2019}, \citealt{smallwood2019}, \citealt{abod2022}, \RR{\citealt{young2023} \& \citealt{ceppi2023}}). It is not feasible to perform such simulations over a large parameter space, so we instead approximate a disk as a geometrically narrow ring with angular momentum equivalent to that of the extended disk it represents. The literature provides a robust theoretical framework for the dynamics of such a ring, with analytical solutions \citep{farago2010,zanazzi2018,martin2019} as well as numerical $n$-body simulations \citep{abod2022,lepp2022} used to translate initial conditions into the ring's dynamical state. For simplicity, we assume that the narrow ring model approximates an extended disk sufficiently well. This is justified if warps and breaks are not important to the disk's evolution, though we discuss these effects in the context of our model.

% \RP{
%     For clarity and consistency with existing literature, it is important to note that disks which undergo retrograde nodal precession (those that are only slightly misaligned) also orbit the binary in the prograde direction, and herein we refer to them as {\em prograde circulating orbits}, or simply as {\em prograde}. Likewise, those disks which undergo prograde nodal precession ($i\sim\pi$) orbit their binary host in the retrograde direction, and we refer to them as {\em retrograde circulating orbits} or simply {\em retrograde}. As noted previously, the nodal precession direction for retrograde circulating orbits switches for sufficiently massive disks. To avoid confusion, we only distinguish between circulating and librating/crescent orbits in this regime. As all nodally precessing circumbinary disks will evolve towards a coplanar orientation, the distinction between prograde and retrograde is not important.
% } \RGM{maybe remove this paragraph, see my suggested sentence at the end of paragraph 4}

@@

For an isotropic \RR{distribution of the angular momentum vector direction (i.e. one where the specific angular momentum $\bm{\ell}$ has no preference for any direction)}, the polar fraction in this limit is
\begin{align}
    f_{{\rm polar}, j \gg 1} &= \int_0^\pi\,\left (1-\frac{2}{\pi}\arcsin{\left (\sqrt{\frac{2}{5}} \csc{i} \right )}\right ) \sin{i}\,di\, , \\
    & = \variable{output/high_j_integral.txt} \, . \nonumber
\end{align}
\RR{This is the value that our calculations in Section \ref{sec:rk} tend to for high $j$ and low $e_\text{b}$}

@@

We determine the dynamical state of the system by tracking the orbital quantities $x=i\,\cos{\Omega}$ and $y=i\,\sin{\Omega}$, where
\begin{equation}
    i = \arccos{\ell_z}
    \label{eq:inclination}
\end{equation}
and
\begin{equation}
    \Omega = \arctantwo (\ell_x, -\ell_y)\, .
    \label{eq:omega}
\end{equation}
% If a disk undergoes \RP{retrograde} nodal precession, then \RR{the path $(x,\,y)$} forms clockwise loops about the origin as the system evolves. Under \RP{prograde} precession, these loops are counter-clockwise.
\RR{If a disk undergoes circulating nodal precession, then the path $(x,\,y)$ forms closed loops about the origin as the system evolves. }
However, any state that would lead to a polar orbit (i.e. libration and crescent patterns) will form closed loops that do not encompass the origin, and these tracks will never cross the line $i\,\sin{\Omega}=0$. The state can be determined by tracking the quadrant \RR{(defined using the standard counter-clockwise convention)} of the point $(i\,\cos{\Omega},~i\,\sin{\Omega})$. \RP{Circulating orbits will traverse all four quadrants.}
A polar orbit that starts in quadrant I or II, however, is restricted to those two quadrants and will alternate between them. The same goes for polar orbits in quadrants III and IV. Therefore, the dynamical state can often be determined by the list of previous quadrant crossings.

@@

For any given system, the dynamical state is determined similarly to the method described in Section \ref{sec:rk}. Note, however, that the $(i\,\cos{\Omega},~i\,\sin{\Omega})$ tracks now contain small structure on the orbital timescale of the outer mass. These structures can mean that a simulation may not make monotonic progress along its track, occasionally backtracking. Such a backtrack can give a false reading of the dynamical state if it occurs across the $x$ or $y$ axis. To mitigate this problem, we integrate each system until it returns to its original position in $(i\,\cos{\Omega},~i\,\sin{\Omega})$ space. If, at this point, the path does not enclose the origin, then it is polar. \RR{Prograde and retrograde circulating orbits can be distinguished by the initial inclination.} Figure \ref{fig:sim_example} shows paths for a set of simulations initialized with $\Omega = \pi/2$.

@@

\begin{figure}
    \begin{centering}
        \includegraphics[width=0.5\textwidth]{figures/compare_massive.pdf}
        \caption{
            MC results (points) for $e_{\rm b}=0.4$, $j=0.66$, and $f_{\rm b}=0.5$ compared to RKF results for the same setup (contour lines).
            \RR{At this value of $j$ all circulating orbits have the same secular behavior regardless of the orbit being prograde or retrograde.}
             The regions interior to the black contour lines are initial conditions which allow polar \RR{librating} orbits according to the RKF method.
            Note that some points that {\sc rebound} classifies as polar fall outside the RKF polar region.
            The MC result gives $f_{\text{polar}} = \variable{output/mc_out.txt}$ while the RKF result gives $f_{\text{polar}} = \variable{output/rk_out.txt}$ with
            numerical error on the order of $\variable{output/rk_err.txt}$. \RR{The point marked ``Unknown'' represents a system whose state could not be determined after $10\,000$ orbits of the third body due to a relatively long dynamical timescale and has been factored into the MC uncertainty.} \RR{The MC simulation results used to generate this figure can be downloaded at \dataset[doi: 10.5281/zenodo.14736363]{\doi{10.5281/zenodo.14736363}}}
        }
        \script{compare_massive.py}
        \label{fig:rkf}
    \end{centering}
\end{figure}

@@

Figure \ref{fig:fpol_j} shows $f_\text{polar}$ as a function of $j$ for various choices of $p_{e_\text{b}}(e_\text{b})$. As expected, the curves converge in the Kozai-Lidov regime, \RR{except for very large $e_{\rm b}=0.99$.} The uniform and thermal distributions both produce polar fractions $>0.5$ for small $j$, with the thermal distribution slightly higher. Except for when $j$ is close to 0, the circularized $e_\text{b} < 0.1$ curve follows the $e_\text{b} = 0.01$ curve very closely, including a maximum at $j=0.5$. This feature is present to some degree in all the low-$j$ curves, and is the result of the fact that, for $e_\text{b}=0$, the critical value for $j$ is 0.5 \citep{abod2022}. \RP{It is around this value of $j$ that crescent orbits begin to appear, and the size and shape of the libration region changes very rapidly. For general $e_\text{b}$, the transition occurs when $j\sim j_\text{cr}$, where $j_\text{cr}$ is the critical value of $j$ \citep[eq. A3]{martin2019}.}

@@

Note that $\Gamma_1$ depends on $e_\text{b}$, and so it is only a constant in the case of a test particle. For $j>0$, we must compute the precession timescale numerically. The time coordinate used in Section \ref{sec:rk} is $\tau \equiv \alpha' t$. $\alpha'$ is defined by \citet[see eq. 3.9]{farago2010}, and we write it as
\begin{equation}
    \alpha' \equiv \frac{3}{4} \Omega_\text{b} \left(\frac{a_\text{b}}{r}\right)^{7/2} f(1-f) \,\Gamma_2(j,f,a_\text{b}/r, e_\text{b}) \, ,
\end{equation}
where we have written the $\sqrt{1+m_\text{r}/M_\text{b}}$ factor as $\Gamma_2$. \RR{$m_\text{r}$ is the mass of the ring. To compute $\Gamma_2$ from our variables requires finding the roots of a cubic equation (see equation (\ref{eq:gamma2-eq})). However, for low $j$ (e.g. $j<10$), $m_\text{r} \ll M_\text{b}$ and $\Gamma_2 \sim 1$.}
